---
title: "Week_6_Practical"
author: "Owen Hughes"
date: "2025-11-11"
output: html_document
---

# This week the goal is to explore analysis of point clusters in a number of different ways

__As always the first steps are to load libraries and data__

```{r}
# Libraries
# install.packages("spatstat")
library(spatstat)
library(here)
library(sp)
library(tmap)
library(sf)
library(tmaptools)
library(tidyverse)
```


```{r}
# Read in the data
LondonBoroughs <- st_read(here::here("Week_6_prac_and_hw/statistical-gis-boundaries-london/statistical-gis-boundaries-london/ESRI/London_Borough_Excluding_MHW.shp"))
```

Let's now inspect the LondonBorough data.
```{r}
# Quick Map
library(stringr)
BoroughMap <- LondonBoroughs %>%
  dplyr::filter(str_detect(GSS_CODE, "^E09"))%>%
  st_transform(., 27700)

qtm(BoroughMap)
```
```{r}
# Summary Stats
summary(BoroughMap)
```
Now let's load the blue plaques data that we will use for the analysis in this practical
```{r}
BluePlaques <- st_read(here::here("Week_6_prac_and_hw/open_plaques.geojson")) %>%
  st_transform(.,27700)

summary(BluePlaques)

```

__Now I want to plot the blue plaques on top of the London Boroughs__

```{r}
tmap_mode("plot")

tm_shape(BoroughMap) +
  tm_polygons(fill_alpha = 0.5) +
tm_shape(BluePlaques) +
  tm_dots(fill = "#57a0d2", size = 0.1)
```

I can see there are a lot of blue plaques in the city of London, Westminster, Camden, Islington?, and fewer once one exits the center of London. 

__There is a plaque point outside of the city, let's clip the points to the city boundary to avoid issues with the analysis__
```{r}
# Remove duplicates
BluePlaques <- distinct(BluePlaques)

# Clip the points
BluePlaquesSub <- BluePlaques[BoroughMap,]
# This code can be used with any topological operator like: intersect or within

tm_shape(BoroughMap) +
  tm_polygons(fill_alpha = 0.5) +
tm_shape(BluePlaquesSub) +
  tm_dots(fill = "#57a0d2", size = 0.1)
```
Great! The errant points have been removed and we're a go for continued analysis.

__Choose a borough for more the sake of the example__
Because of the amount of points in the data set for the whole city, choosing a borough will help speed up the process. I am going to choose Lambeth because I like Brixton a lot.

```{r}
# Extract Lambeth
Lambeth <- BoroughMap %>%
  filter(., NAME=="Lambeth")

# Check to verify it's correct
tm_shape(Lambeth)+
  tm_polygons(fill = "#EAA221", fill_alpha = 0.5)
```

Now I want to clip the plaques in Lambeth to the map above and check
```{r}
# Clip to Lambeth
BluePlaquesLam <- BluePlaques[Lambeth,]

# Map to check
tm_shape(Lambeth) +
  tm_polygons(fill = "#EAA221", fill_alpha = 0.5)+
tm_shape(BluePlaquesLam) +
  tm_dots(fill = "#57a0d2", size = 0.3) # slightly larger dots are easier to see

```

__Final setup for point cloud analysis__

Create a spatstat observation window
```{r}
#now set a window as the borough boundary
window <- as.owin(Lambeth)
plot(window)
```
Spatstat does not work directly with sf objects so let's convert the data to something that spatstat can read properly. 
```{r}
#create a sp object
BluePlaquesLam<- BluePlaquesLam %>%
  as(., 'Spatial')
#create a ppp object
BluePlaquesLam.ppp <- ppp(x=BluePlaquesLam@coords[,1], # grabs lat + long from geometry column
                          y=BluePlaquesLam@coords[,2],
                          window=window)
```


Inspect the new spatstat object
```{r}
BluePlaquesLam.ppp %>%
  plot(.,pch=16,cex=0.5, 
       main="Blue Plaques Lambeth")
```

__Point Pattern Analysis__
