---
title: "Week_8_9_prac"
output: html_document
---

# Practical Week 8 & 9

## Load in Data

```{r}
library(tidyverse)
library(tmap)
library(plotly)
library(broom)
#install.packages("mapview")
library(mapview)
library(sf)
library(sp)
library(spdep)
#install.packages('car')
library(car)
library(fs)
library(janitor)
```

```{r}
# Read in data
download.file("https://data.london.gov.uk/download/statistical-gis-boundary-files-london/9ba8c833-6370-4b11-abdc-314aa020d5e0/statistical-gis-boundaries-london.zip",
              destfile = "statistical-gis-boundaries-london.zip")
```

```{r}
# Get the file and extract it
library(fs)
listfiles<-dir_info(here::here("Week_8_prac")) %>%
  dplyr::filter(str_detect(path, ".zip")) %>%
  dplyr::select(path)%>%
  pull()%>%
  #print out the .gz file
  print()%>%
  as.character()%>%
  utils::unzip(exdir=here::here("Week_8_prac"))
```

```{r}
#look what is inside the zip

Londonwards<-fs::dir_info(here::here("Week_8_prac", 
                                 "statistical-gis-boundaries-london", 
                                 "ESRI"))%>%
  #$ means exact match
  dplyr::filter(str_detect(path, 
                           "London_Ward_CityMerged.shp$"))%>%
  dplyr::select(path)%>%
  dplyr::pull()%>%
  #read in the file in
  sf::st_read()
```

```{r}
#check the data
qtm(Londonwards)
```

```{r}
#read in some attribute data
LondonWardProfiles <- read_csv("https://data.london.gov.uk/download/f33fb38c-cb37-48e3-8298-84c0d3cc5a6c/772d2d64-e8c6-46cb-86f9-e52b4c7851bc/ward-profiles-excel-version.csv", 
                               col_names = TRUE, 
                               locale = locale(encoding = 'Latin1'))
```

```{r}
#check all of the columns have been read in correctly
Datatypelist <- LondonWardProfiles %>% 
  summarise_all(class) %>%
  pivot_longer(everything(), 
               names_to="All_variables", 
               values_to="Variable_class")

Datatypelist
```

#### Cleaning the data as you read it in

```{r}
#We can use readr to deal with the issues in this dataset - which are to do with text values being stored in columns containing numeric values

#read in some data - couple of things here. Read in specifying a load of likely 'n/a' values, also specify Latin1 as encoding as there is a pound sign (£) in one of the column headers - just to make things fun!

LondonWardProfiles <- read_csv("https://data.london.gov.uk/download/ward-profiles-and-atlas/772d2d64-e8c6-46cb-86f9-e52b4c7851bc/ward-profiles-excel-version.csv", 
                               na = c("", "NA", "n/a"), 
                               locale = locale(encoding = 'Latin1'), 
                               col_names = TRUE)
```

```{r}
#check all of the columns have been read in correctly
Datatypelist <- LondonWardProfiles %>% 
  summarise_all(class) %>%
  pivot_longer(everything(), 
               names_to="All_variables", 
               values_to="Variable_class")

Datatypelist
```

```{r}
#merge boundaries and data
LonWardProfiles <- Londonwards%>%
  left_join(.,
            LondonWardProfiles, 
            by = c("GSS_CODE" = "New code"))


#let's map our dependent variable to see if the join has worked:
tmap_mode("plot")
qtm(LonWardProfiles, 
    fill = "Average GCSE capped point scores - 2014", 
    fill.palette = "oranges")
```

#### Additional Data

```{r}
#might be a good idea to see where the secondary schools are in London too
london_schools <- read_csv("https://data.london.gov.uk/download/146392df-b051-42ad-b8ec-454e440f0f8b/57046151-39a0-45d9-8dc0-27ea7fd02de8/all_schools_xy_2016.csv")

#from the coordinate values stored in the x and y columns, which look like they are latitude and longitude values, create a new points dataset
lon_schools_sf <- st_as_sf(london_schools, 
                           coords = c("x","y"), 
                           crs = 4326)

lond_sec_schools_sf <- lon_schools_sf %>%
  filter(PHASE=="Secondary")

tmap_mode("plot")
qtm(lond_sec_schools_sf, size=0.2)
```

## Analyzing GCSE exam performance - testing a research hypothesis

To explore the factors that might influence GCSE exam performance in London, we are going to run a series of different regression models. A regression model is simply the expression of a linear relationship between our outcome variable (Average GCSE score in each Ward in London) and another variable or several variables that might explain this outcome.

### Research Question and Hypothesis

Examining the spatial distribution of GSCE point scores in the map above, it is clear that there is variation across the city. My research question is:

**What are the factors that might lead to variation in Average GCSE point scores across the city?**

My research hypothesis that I am going to test is that there are other observable factors occurring in Wards in London that might affect the average GCSE scores of students living in those areas.

In inferential statistics, we cannot definitively prove a hypothesis is true, but we can seek to *disprove* that there is absolutely nothing of interest occurring or no association between variables. The *null hypothesis* that I am going to test empirically with some models is that there is *no* relationship between exam scores and other observed variables across London.

### Regression Basics

```{r}
q <- qplot(x = `Unauthorised Absence in All Schools (%) - 2013`, 
           y = `Average GCSE capped point scores - 2014`, 
           data=LonWardProfiles)
```

```{r}
#plot with a regression line - note, I've added some jitter here as the x-scale is rounded
q + stat_smooth(method="lm", se=FALSE, size=1) + 
  geom_jitter()
```

### Running a Regression Model in R

```{r}
#run the linear regression model and store its outputs in an object called model1
Regressiondata<- LonWardProfiles%>%
  clean_names()%>%
  dplyr::select(average_gcse_capped_point_scores_2014, 
                unauthorised_absence_in_all_schools_percent_2013)

#now model
model1 <- Regressiondata %>%
  lm(average_gcse_capped_point_scores_2014 ~
               unauthorised_absence_in_all_schools_percent_2013,
     data=.)
```

```{r}
#show the summary of those outputs
summary(model1)
```

### Broom

```{r}
library(broom)
tidy(model1)
```

We can also use `glance()` from `broom` to get a bit more summary information, such as r2 and the adjusted r-squared value

```{r}
glance(model1)
```

But wait? Didn’t we try to model our GCSE values based on our unauthorized absence variable? Can we see those predictions for each point, yes, yes we can…with the `tidypredict_to_column()` function from `tidypredict`, which adds the `fit` column in the following code.

```{r}
#install.packages('tidypredict')
library(tidypredict)
Regressiondata %>%
  tidypredict_to_column(model1)
```

### Tidymodels

Before we move on it’s worth pointing out that a new iteration of modelling is being developed through [`tidymodels`](https://www.tidymodels.org/)…the benefit of this is that we can easily change the modelling method or as they term it…`engine`…(e.g. to [RandomForest](https://www.rebeccabarter.com/blog/2020-03-25_machine_learning/))

```{r}
#install.packages("tidymodels")
library(tidymodels)
```

```{r}
# set the model
lm_mod <- linear_reg()

# fit the model
lm_fit <- 
  lm_mod %>% 
  fit(average_gcse_capped_point_scores_2014 ~
               unauthorised_absence_in_all_schools_percent_2013,
     data=Regressiondata)

# we cover tidy and glance in a minute...
tidy(lm_fit)
```

```{r}
glance(lm_fit)
```

### Bootstrap Resampling. Cont.

If we only fit our model once, how can we be confident about that estimate? Bootstrap resampling is where we take the original dataset and select random data points from within it, but in order to keep it the same size as the original dataset some records are duplicated. This is known as bootstrap resampling by replacement. We used to briefly cover this within this practical but have recently removed it. If you wish to explore it then consult the [bootstrap resampling section from previous years](https://andrewmaclachlan.github.io/CASA0005repo_20202021/gwr-and-spatially-lagged-regression.html#bootstrap-resampling), but this is not a requirement and only for interest.

### Variables

#### How do I select variables

**Either with logic and reasoning (e.g. in the exam),** with the support of academic literature or methods such as [best subset regression, k-fold cross validation](http://www.sthda.com/english/articles/37-model-selection-essentials-in-r/155-best-subsets-regression-essentials-in-r/) or [gradient descent](https://mlu-explain.github.io/linear-regression/). It does not form part of this module. Other regression methods such as [Ridge, LASSO and elastic net regression](https://andrewmaclachlan.github.io/CASA0005repo_20202021/advanced-r-maup-and-more-regression.html#advanced-regression) can reduce the influence of variables that are not useful in the model. But again, this is beyond the scope of this module.

### Assumptions Underpinning Linear Regression

#### Assumption 1 - There is a linear relationship between the dependent and independent variables

Do this by plotting the distribution (histogram) of the variables being tested

```{r}
ggplot(LonWardProfiles, aes(x=`Average GCSE capped point scores - 2014`)) + 
  geom_histogram(aes(y = ..density..),
                 binwidth = 5) + 
  geom_density(colour="red", 
               size=1, 
               adjust=1)
```

```{r}
ggplot(LonWardProfiles, aes(x=`Unauthorised Absence in All Schools (%) - 2013`)) +
  geom_histogram(aes(y = ..density..),
                 binwidth = 0.1) + 
  geom_density(colour="red",
               size=1, 
               adjust=1)
```

The two plots above are sufficiently normal, let's compare to a variable that isn't so normal.

```{r}
library(ggplot2)

# from 21/10 there is an error on the website with 
# median_house_price_2014 being called median_house_price<c2>2014
# this was corrected around 23/11 but can be corrected with rename..

LonWardProfiles <- LonWardProfiles %>%
  #try removing this line to see if it works...
  dplyr::rename(median_house_price_2014 =`Median House Price (£) - 2014`)%>%
  janitor::clean_names()

ggplot(LonWardProfiles, aes(x=median_house_price_2014)) + 
  geom_histogram()
```

This is what we get if we plot the raw house prices vs the GCSE scores

```{r}
qplot(x = median_house_price_2014, 
      y = average_gcse_capped_point_scores_2014, 
      data=LonWardProfiles)
```

Variable Transformations

Sometimes, like the case of the house prices, one can take the log of the variable to change its distribution to be normal.

```{r}
ggplot(LonWardProfiles, aes(x=log(median_house_price_2014))) + 
  geom_histogram()
```

As can be seen, the distribution is more normal than it was, but still not perfect.

There are several other possible transformations, the symbox function allows us to investigate several at the same time.

```{r}
symbox(~median_house_price_2014, 
       LonWardProfiles, 
       na.rm=T,
       powers=seq(-3,3,by=.5))
```

It looks like a power of -1 should be the most normal

```{r}
ggplot(LonWardProfiles, aes(x=(median_house_price_2014)^-1)) + 
  geom_histogram()
```

Here is the comparison scatter plot of the power = -1 with the GCSE scores

```{r}
qplot(x = (median_house_price_2014)^-1, 
      y = average_gcse_capped_point_scores_2014,
      data=LonWardProfiles)
```

And here is the scatter plot with the log transformations

```{r}
qplot(x = log(median_house_price_2014), 
      y = average_gcse_capped_point_scores_2014, 
      data=LonWardProfiles)
```

These seem to have the opposite relationship with GCSE scores, so the way we analyze these relationships depends on the type of transformation.

#### Assumption 2 - The residuals in your model should be normally distributed

This assumption is easy to check. When we ran our Model1 earlier, one of the outputs stored in our Model 1 object is the residual value for each case (Ward) in your data set. We can access these values using `augment()` from `broom` which will add model output to the original GCSE data…

We can plot these as a histogram and see if there is a normal distribution:

```{r}
#save the residuals into your dataframe

model_data <- model1 %>%
  augment(., Regressiondata)

#plot residuals
model_data%>%
dplyr::select(.resid)%>%
  pull()%>%
  qplot()+ 
  geom_histogram() 
```

Looks pretty normal to me

#### Assumption 3 - No Multicolinearity in the independent variables

Now, the regression model we have be experimenting with so far is a simple bivariate (two variable) model. One of the nice things about regression modelling is while we can only easily visualise linear relationships in a two (or maximum 3) dimension scatter plot, mathematically, we can have as many dimensions / variables as we like.

As such, we could extend model 1 into a **multiple regression** model by adding some more explanatory variables that we think could affect GSCE scores. Let’s try the log or \^-1 transformed house price variable from earlier (the rationale being that higher house prices indicate more affluence and therefore, potentially, more engagement with education):

```{r}
Regressiondata2<- LonWardProfiles%>%
  clean_names()%>%
  dplyr::select(average_gcse_capped_point_scores_2014,
         unauthorised_absence_in_all_schools_percent_2013,
         median_house_price_2014)

model2 <- lm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
               log(median_house_price_2014), data = Regressiondata2)

#show the summary of those outputs
tidy(model2)
```

```{r}
glance(model2)
```

```{r}
#and for future use, write the residuals out
model_data2 <- model2 %>%
  augment(., Regressiondata2)

# also add them to the shapelayer
LonWardProfiles <- LonWardProfiles %>%
  mutate(model2resids = residuals(model2))
```

We can see that both variables are significant, but let's check if they are co-linear

```{r}
#install.packages('corrr')
library(corrr)

Correlation <- LonWardProfiles %>%
  st_drop_geometry()%>%
  dplyr::select(average_gcse_capped_point_scores_2014,
         unauthorised_absence_in_all_schools_percent_2013,
         median_house_price_2014) %>%
  mutate(median_house_price_2014 =log(median_house_price_2014))%>%
    correlate() %>%
  # just focus on GCSE and house prices
  focus(-average_gcse_capped_point_scores_2014, mirror = TRUE) 


#visualise the correlation matrix
rplot(Correlation)
```

We can see that these two have relatively low colinearity

**Variance Inflation Factor** (VIF) is another way to check multicolinearity

```{r}
vif(model2)
```

If we wanted to check this for several more variables we could make a correlation matrix.

```{r}
position <- c(10:74)

Correlation_all<- LonWardProfiles %>%
  st_drop_geometry()%>%
  dplyr::select(position)%>%
    correlate()
```

```{r}
rplot(Correlation_all)
```

#### Assumption 4 - Homoscedasticity

Homoscedasticity means that the errors/residuals in the model exhibit constant / homogenous variance, if they don’t, then we would say that there is hetroscedasticity present. Why does this matter? Andy Field does a much better job of explaining this in [discovering statistics](https://www.discoveringstatistics.com/tag/homoscedasticity/) — but essentially, if your errors do not have constant variance, then your parameter estimates could be wrong, as could the estimates of their significance.

The best way to check for homo/hetroscedasticity is to plot the residuals in the model against the predicted values. We are looking for a cloud of points with no apparent patterning to them

```{r}
#print some model diagnositcs. 
par(mfrow=c(2,2))    #plot to 2 by 2 array
plot(model2)
```

```{r}
library(performance)

check_model(model2, check="all")
```

#### Assumption 5 - Independence of Errors

If you are running a regression model on data that do not have explicit space or time dimensions, then the standard test for autocorrelation would be the Durbin-Watson test.

This tests whether residuals are correlated and produces a summary statistic that ranges between 0 and 4, with 2 signifiying no autocorrelation. A value greater than 2 suggesting negative autocorrelation and and value of less than 2 indicating postitve autocorrelation.

In his [excellent text book](https://www.discoveringstatistics.com/books/discovering-statistics-using-r/), Andy Field suggests that you should be concerned with Durbin-Watson test statistics \<1 or \>3. So let’s see:

```{r}
#run durbin-watson test
DW <- durbinWatsonTest(model2)
tidy(DW)
```

As you can see, the DW statistics for our model is 1.61, so some indication of autocorrelation, but perhaps nothing to worry about.

**HOWEVER**

We are using spatially referenced data and as such we should check for *spatial*-autocorrelation.

The first test we should carry out is to map the residuals to see if there are any apparent obvious patterns:

```{r}
#now plot the residuals
tmap_mode("view")
#qtm(LonWardProfiles, fill = "model1_resids")


tm_shape(LonWardProfiles) +
  tm_polygons("model2resids",
               fill.scale = tm_scale_intervals(
                values="brewer.rd_yl_bu",
                n=5,
                style="jenks")) +
  tm_shape(lond_sec_schools_sf) + 
  tm_dots(col = "TYPE")
```

Looking at the map above, there look to be some blue areas next to other blue areas and some red/orange areas next to other red/orange areas. This suggests that there could well be some spatial autocorrelation biasing our model, but can we test for spatial autocorrelation more systematically?

Yes - and some of you will remember this from the practical two weeks ago. We can calculate a number of different statistics to check for spatial autocorrelation - the most common of these being Moran’s I.

```{r}
#calculate the centroids of all Wards in London
coordsW <- LonWardProfiles%>%
  st_centroid()%>%
  st_geometry()

plot(coordsW)
```

```{r}
#Now we need to generate a spatial weights matrix 
#(remember from the lecture a couple of weeks ago). 
#We'll start with a simple binary matrix of queen's case neighbours

LWard_nb <- LonWardProfiles %>%
  poly2nb(., queen=T)

#or nearest neighbours
knn_wards <-coordsW %>%
  knearneigh(., k=4)

LWard_knn <- knn_wards %>%
  knn2nb()

#plot them
plot(LWard_nb, st_geometry(coordsW), col="red")
```

```{r}
plot(LWard_knn, st_geometry(coordsW), col="blue")
```

```{r}
#create a spatial weights matrix object from these weights

Lward.queens_weight <- LWard_nb %>%
  nb2listw(., style="W")

Lward.knn_4_weight <- LWard_knn %>%
  nb2listw(., style="W")
```

Now run a moran’s I test on the residuals, first using queens neighbours

```{r}
Queen <- LonWardProfiles %>%
  st_drop_geometry()%>%
  dplyr::select(model2resids)%>%
  pull()%>%
  moran.test(., Lward.queens_weight)%>%
  tidy()
```

Then nearest k-nearest neighbors

```{r}
Nearest_neighbour <- LonWardProfiles %>%
  st_drop_geometry()%>%
  dplyr::select(model2resids)%>%
  pull()%>%
  moran.test(., Lward.knn_4_weight)%>%
  tidy()

Queen
```

```{r}
Nearest_neighbour
```

Observing the Moran’s I statistic for both Queen’s case neighbors and k-nearest neighbors of 4, we can see that the Moran’s I statistic is somewhere between 0.27 and 0.29. Remembering that Moran’s I ranges from between -1 and +1 (0 indicating no spatial autocorrelation) we can conclude that there is some weak to moderate spatial autocorrelation in our residuals.

This means that despite passing most of the assumptions of linear regression, we could have a situation here where the presence of some spatial autocorrelation could be leading to biased estimates of our parameters and significance values.

### Spatial Regression Models

#### Lag Dependent Model

```{r}
#Original Model
model2 <- lm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
               log(median_house_price_2014), data = LonWardProfiles)

tidy(model2)
```

Queen's Case Lag –\> spatially lagged

```{r}
#install.packages('spatialreg')
library(spatialreg)

slag_dv_model2_queen <- lagsarlm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
               log(median_house_price_2014), 
               data = LonWardProfiles, 
               nb2listw(LWard_nb, style="C"), 
               method = "eigen")

#what do the outputs show?
tidy(slag_dv_model2_queen)
```

```{r}
#glance() gives model stats but this need something produced from a linear model
#here we have used lagsarlm()
glance(slag_dv_model2_queen)
```

```{r}
t<-summary(slag_dv_model2_queen)

sum(t$residuals)
```

```{r}
#install.packages('lmtest')
library(lmtest)
lrtest(slag_dv_model2_queen, model2)
```

```{r}
# weight list is just the code from the lagsarlm
weight_list<-nb2listw(LWard_knn, style="C")

imp <- impacts(slag_dv_model2_queen, listw=weight_list)

imp
```

Now it is appropriate to compare these coefficients to the OLS outputs…**however** if you have a very large matrix this might not work, instead a sparse matrix that uses approximation methods (see Solymosi and Medina (2022) and within that resource, Lesage and Pace 2009). This is beyond the scope of the content here, but essentially this makes the method faster on larger data…but only row standardised is permitted here…

```{r}
slag_dv_model2_queen_row <- lagsarlm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
               log(median_house_price_2014), 
               data = LonWardProfiles, 
               nb2listw(LWard_nb, style="W"), 
               method = "eigen")


W <- as(weight_list, "CsparseMatrix")

trMatc <- trW(W, type="mult")
trMC <- trW(W, type="MC")

imp2 <- impacts(slag_dv_model2_queen_row, tr=trMatc, R=200)

imp3 <- impacts(slag_dv_model2_queen_row, tr=trMC, R=200)

imp2
```

```{r}
imp3
```

We can also get the p-values (where an R is set, this is the number of simulations to use)…from the sparse computation

```{r}
sum <- summary(imp2,  zstats=TRUE, short=TRUE)

sum
```

KNN Case Lag –\> queens case

```{r}
#run a spatially lagged regression model
slag_dv_model2_knn4 <- lagsarlm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
               log(median_house_price_2014), 
               data = LonWardProfiles, 
               nb2listw(LWard_knn, 
                        style="C"), 
               method = "eigen")

#what do the outputs show?
tidy(slag_dv_model2_knn4)
```

We can also now check that the residuals from the spatially lagged model are now no-longer exhibiting spatial autocorrelation:

```{r}
#write out the residuals

LonWardProfiles <- LonWardProfiles %>%
  mutate(slag_dv_model2_knn_resids = residuals(slag_dv_model2_knn4))

KNN4Moran <- LonWardProfiles %>%
  st_drop_geometry()%>%
  dplyr::select(slag_dv_model2_knn_resids)%>%
  pull()%>%
  moran.test(., Lward.knn_4_weight)%>%
  tidy()

KNN4Moran
```

#### The Spatial Error Mode

```{r}
sem_model1 <- errorsarlm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
               log(median_house_price_2014), 
               data = LonWardProfiles,
               nb2listw(LWard_knn, style="C"), 
               method = "eigen")

tidy(sem_model1)
```

#### Key advice

Usually you might run a **lag** model when you have an idea of what is causing the spatial autocorrelation in the dependent variable and an **error** model when you aren’t sure what might be missing. However, you can also use a more scientific method - the Lagrange Multiplier test.

But! recall from a few weeks ago when I made a big deal of type of standardisation for the spatial weight matrix? This test **expects** row standardisation.

The Lagrange multiple tests are within the function `lm.LMtests` from the package `spdep`:

-   `LMerr` is the spatial error model test

-   `LMlag` is the lagged test

With each also having a robust form, being robust to insensitivities to changes (e.g. outliers, non-normality).

```{r}
library(spdep)

Lward.queens_weight_ROW <- LWard_nb %>%
  nb2listw(., style="W")

lm.LMtests(model2, Lward.queens_weight_ROW, test = c("LMerr","LMlag","RLMerr","RLMlag","SARMA"))
```

#### Which model to use

Usually you will run OLS regression first then look for spatial autocorrelation of the residuals (Moran’s I).

Once at this stage you need to make a decision about the model:

-   Is it a global model (error / lag) or a local model (GWR)?

-   Can a single model (error/lag) be fitted to the study area?

-   Is the spatial autocorrelation a problem (error) or showing local trends (GWR)?

Of course you could do a OLS, spatial lag and GWR as long as they all contribute something to your analysis.

#### More data

We will now read in some extra data which we will use shortly

```{r}
extradata <- read_csv("https://www.dropbox.com/s/qay9q1jwpffxcqj/LondonAdditionalDataFixed.csv?raw=1")

#add the extra data too
LonWardProfiles <- LonWardProfiles%>%
  left_join(., 
            extradata, 
            by = c("gss_code" = "Wardcode"))%>%
  clean_names()

#print some of the column names
LonWardProfiles%>%
  names()%>%
  tail(., n=10)
```

####  Extending your regression model - Dummy Variables

```{r}
p <- ggplot(LonWardProfiles, 
            aes(x=unauth_absence_schools11, 
                y=average_gcse_capped_point_scores_2014))
p + geom_point(aes(colour = inner_outer)) 
```

```{r}
#first, let's make sure R is reading our InnerOuter variable as a factor
#see what it is at the moment...

Datatypelist <- LonWardProfiles %>%
  st_drop_geometry%>% 
#summarise_all only works with .tbl now (not sf) so we   drop geometry to check  
  summarise_all(class)%>%
  pivot_longer(everything(), 
             names_to="All_variables", 
             values_to="Variable_class")

Datatypelist
```

```{r}
# change to factor

LonWardProfiles<- LonWardProfiles %>%
  mutate(inner_outer=as.factor(inner_outer))

#now run the model
model3 <- lm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
               log(median_house_price_2014) + 
               inner_outer, 
             data = LonWardProfiles)
 
tidy(model3)
```

```{r}
contrasts(LonWardProfiles$inner_outer)
```

Here's how to change the control dummy variable

```{r}
LonWardProfiles <- LonWardProfiles %>%
  mutate(inner_outer = relevel(inner_outer, 
                               ref="Outer"))

model3 <- lm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
               log(median_house_price_2014) + 
               inner_outer,
             data = LonWardProfiles)

tidy(model3)
```

## Task 3 - Spatial Non-stationarity and Geographically Weighted Regression Models (GWR)

Here is a complete model

```{r}
#select some variables from the data file
myvars <- LonWardProfiles %>%
  dplyr::select(average_gcse_capped_point_scores_2014,
         unauthorised_absence_in_all_schools_percent_2013,
         median_house_price_2014,
         rate_of_job_seekers_allowance_jsa_claimants_2015,
         percent_with_level_4_qualifications_and_above_2011,
         inner_outer)

#check their correlations are OK
Correlation_myvars <- myvars %>%
  st_drop_geometry()%>%
  dplyr::select(-inner_outer)%>%
  correlate()

#run a final OLS model
model_final <- lm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
                    log(median_house_price_2014) + 
                    inner_outer + 
                    rate_of_job_seekers_allowance_jsa_claimants_2015 +
                    percent_with_level_4_qualifications_and_above_2011, 
                  data = myvars)

tidy(model_final)
```

```{r}
LonWardProfiles <- LonWardProfiles %>%
  mutate(model_final_res = residuals(model_final))

par(mfrow=c(2,2))
plot(model_final)
```

```{r}
qtm(LonWardProfiles, fill = "model_final_res")
```

```{r}
final_model_Moran <- LonWardProfiles %>%
  st_drop_geometry()%>%
  dplyr::select(model_final_res)%>%
  pull()%>%
  moran.test(., Lward.knn_4_weight)%>%
  tidy()

final_model_Moran
```

```{r}
#install.packages('spgwr')
library(spgwr)

coordsW2 <- st_coordinates(coordsW)

LonWardProfiles2 <- cbind(LonWardProfiles,coordsW2)

GWRbandwidth <- gwr.sel(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
                    log(median_house_price_2014) + 
                    inner_outer + 
                    rate_of_job_seekers_allowance_jsa_claimants_2015 +
                    percent_with_level_4_qualifications_and_above_2011, 
                    data = LonWardProfiles2, 
                    coords=cbind(LonWardProfiles2$X, 
                                 LonWardProfiles2$Y),
                  adapt=T)
```
