---
title: "Week_8_9_prac"
output: html_document
---

# Practical Week 8 & 9

## Load in Data

```{r}
library(tidyverse)
library(tmap)
library(plotly)
library(broom)
#install.packages("mapview")
library(mapview)
library(sf)
library(sp)
library(spdep)
#install.packages('car')
library(car)
library(fs)
library(janitor)
```

```{r}
# Read in data
download.file("https://data.london.gov.uk/download/statistical-gis-boundary-files-london/9ba8c833-6370-4b11-abdc-314aa020d5e0/statistical-gis-boundaries-london.zip",
              destfile = "statistical-gis-boundaries-london.zip")
```

```{r}
# Get the file and extract it
library(fs)
listfiles<-dir_info(here::here("Week_8_prac")) %>%
  dplyr::filter(str_detect(path, ".zip")) %>%
  dplyr::select(path)%>%
  pull()%>%
  #print out the .gz file
  print()%>%
  as.character()%>%
  utils::unzip(exdir=here::here("Week_8_prac"))
```

```{r}
#look what is inside the zip

Londonwards<-fs::dir_info(here::here("Week_8_prac", 
                                 "statistical-gis-boundaries-london", 
                                 "ESRI"))%>%
  #$ means exact match
  dplyr::filter(str_detect(path, 
                           "London_Ward_CityMerged.shp$"))%>%
  dplyr::select(path)%>%
  dplyr::pull()%>%
  #read in the file in
  sf::st_read()
```

```{r}
#check the data
qtm(Londonwards)
```

```{r}
#read in some attribute data
LondonWardProfiles <- read_csv("https://data.london.gov.uk/download/f33fb38c-cb37-48e3-8298-84c0d3cc5a6c/772d2d64-e8c6-46cb-86f9-e52b4c7851bc/ward-profiles-excel-version.csv", 
                               col_names = TRUE, 
                               locale = locale(encoding = 'Latin1'))
```

```{r}
#check all of the columns have been read in correctly
Datatypelist <- LondonWardProfiles %>% 
  summarise_all(class) %>%
  pivot_longer(everything(), 
               names_to="All_variables", 
               values_to="Variable_class")

Datatypelist
```

#### Cleaning the data as you read it in

```{r}
#We can use readr to deal with the issues in this dataset - which are to do with text values being stored in columns containing numeric values

#read in some data - couple of things here. Read in specifying a load of likely 'n/a' values, also specify Latin1 as encoding as there is a pound sign (£) in one of the column headers - just to make things fun!

LondonWardProfiles <- read_csv("https://data.london.gov.uk/download/ward-profiles-and-atlas/772d2d64-e8c6-46cb-86f9-e52b4c7851bc/ward-profiles-excel-version.csv", 
                               na = c("", "NA", "n/a"), 
                               locale = locale(encoding = 'Latin1'), 
                               col_names = TRUE)
```

```{r}
#check all of the columns have been read in correctly
Datatypelist <- LondonWardProfiles %>% 
  summarise_all(class) %>%
  pivot_longer(everything(), 
               names_to="All_variables", 
               values_to="Variable_class")

Datatypelist
```

```{r}
#merge boundaries and data
LonWardProfiles <- Londonwards%>%
  left_join(.,
            LondonWardProfiles, 
            by = c("GSS_CODE" = "New code"))


#let's map our dependent variable to see if the join has worked:
tmap_mode("plot")
qtm(LonWardProfiles, 
    fill = "Average GCSE capped point scores - 2014", 
    fill.palette = "oranges")
```

#### Additional Data

```{r}
#might be a good idea to see where the secondary schools are in London too
london_schools <- read_csv("https://data.london.gov.uk/download/146392df-b051-42ad-b8ec-454e440f0f8b/57046151-39a0-45d9-8dc0-27ea7fd02de8/all_schools_xy_2016.csv")

#from the coordinate values stored in the x and y columns, which look like they are latitude and longitude values, create a new points dataset
lon_schools_sf <- st_as_sf(london_schools, 
                           coords = c("x","y"), 
                           crs = 4326)

lond_sec_schools_sf <- lon_schools_sf %>%
  filter(PHASE=="Secondary")

tmap_mode("plot")
qtm(lond_sec_schools_sf, size=0.2)
```

## Analysing GCSE exam performance - testing a research hypothesis

To explore the factors that might influence GCSE exam performance in London, we are going to run a series of different regression models. A regression model is simply the expression of a linear relationship between our outcome variable (Average GCSE score in each Ward in London) and another variable or several variables that might explain this outcome.

### Research Question and Hypothesis

Examining the spatial distribution of GSCE point scores in the map above, it is clear that there is variation across the city. My research question is:

**What are the factors that might lead to variation in Average GCSE point scores across the city?**

My research hypothesis that I am going to test is that there are other observable factors occurring in Wards in London that might affect the average GCSE scores of students living in those areas.

In inferential statistics, we cannot definitively prove a hypothesis is true, but we can seek to *disprove* that there is absolutely nothing of interest occurring or no association between variables. The *null hypothesis* that I am going to test empirically with some models is that there is *no* relationship between exam scores and other observed variables across London.

### Regression Basics

```{r}
q <- qplot(x = `Unauthorised Absence in All Schools (%) - 2013`, 
           y = `Average GCSE capped point scores - 2014`, 
           data=LonWardProfiles)
```

```{r}
#plot with a regression line - note, I've added some jitter here as the x-scale is rounded
q + stat_smooth(method="lm", se=FALSE, size=1) + 
  geom_jitter()
```

### Running a Regression Model in R

```{r}
#run the linear regression model and store its outputs in an object called model1
Regressiondata<- LonWardProfiles%>%
  clean_names()%>%
  dplyr::select(average_gcse_capped_point_scores_2014, 
                unauthorised_absence_in_all_schools_percent_2013)

#now model
model1 <- Regressiondata %>%
  lm(average_gcse_capped_point_scores_2014 ~
               unauthorised_absence_in_all_schools_percent_2013,
     data=.)
```

```{r}
#show the summary of those outputs
summary(model1)
```

### Broom

```{r}
library(broom)
tidy(model1)
```

We can also use `glance()` from `broom` to get a bit more summary information, such as r2 and the adjusted r-squared value

```{r}
glance(model1)
```

But wait? Didn’t we try to model our GCSE values based on our unauthorised absence variable? Can we see those predictions for each point, yes, yes we can…with the `tidypredict_to_column()` function from `tidypredict`, which adds the `fit` column in the following code.

```{r}
#install.packages('tidypredict')
library(tidypredict)
Regressiondata %>%
  tidypredict_to_column(model1)
```

### Tidymodels

Before we move on it’s worth pointing out that a new iteration of modelling is being developed through [`tidymodels`](https://www.tidymodels.org/)…the benefit of this is that we can easily change the modelling method or as they term it…`engine`…(e.g. to [RandomForest](https://www.rebeccabarter.com/blog/2020-03-25_machine_learning/))

```{r}
#install.packages("tidymodels")
library(tidymodels)
```

```{r}
# set the model
lm_mod <- linear_reg()

# fit the model
lm_fit <- 
  lm_mod %>% 
  fit(average_gcse_capped_point_scores_2014 ~
               unauthorised_absence_in_all_schools_percent_2013,
     data=Regressiondata)

# we cover tidy and glance in a minute...
tidy(lm_fit)
```

```{r}
glance(lm_fit)
```

### Bootstrap Resampling. Cont. 
